The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-05-26 brains-conclusion.tex
chore/todos: note all github issues in the main text

% !TEX root = brainscopycut.tex
\section{Concluding remarks}\label{sec:conclusion}
We aimed to contribute to the empirical understanding of representation transformation processes %\marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?}
 by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content. To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
In more detail, we describe the joint properties of the substituted and substituting terms in the reformulation by individuals of a specific type of utterances (quotations). %--- in this sense we also diverge from psycholinguistic experiments that focus on ease of recall.

>>>>

% !TEX root = brainscopycut.tex
\section{Concluding remarks}\label{sec:conclusion}

\todo{\#14: link to introduction discussion, mentioning this can be a model system}

\todo{\#15: relate to missed literature}

We aimed to contribute to the empirical understanding of representation transformation processes %\marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?}
 by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content. To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
In more detail, we describe the joint properties of the substituted and substituting terms in the reformulation by individuals of a specific type of utterances (quotations). %--- in this sense we also diverge from psycholinguistic experiments that focus on ease of recall.

\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}

### 2016-05-26 brains-introduction.tex
chore/todos: note all github issues in the main text

@@

>>>>

\todo{
\#14: Rework to
(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
(2) present current problem of testing theories to incorporate those discoveries,
(3) present big data literature which can bring measurement tools,
(4) introduce the question of cultural attractors,
(5) to be studied through quotes on the web: it's the proposal of a novel approach, we'll be describing and cross-checking with existing literature (vs. predicting predict, since there are too many possible things to predict)
}

@@

\todo{
\#15: add
(1) sentence recall (Potter \& Lombardi),
(2) false memories (Deese),
(3) subjective organization (Tulving, Zaromb),
(4) working memory and attention (Jefferies),
(5) iterated learning (Kirby)
}

### 2016-05-26 brains-protocol.tex
chore/todos: note all github issues in the main text

{We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.} To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.

@@

@@

@@

@@

@@

\medskip
Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Fig.~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.  In practice, from the 2.43m initial occurrences spread into 45,749 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 6,172 real substitutions obeying to this model. From these substitutions we remove those featuring stop words, minor spelling changes (e.g. center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 1,051 valid substitutions.

@@

@@

%\begin{figure*}[!th]
%    \centering
%    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
%    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
%    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
%    \label{fig:feature-susceptibilities}
%\end{figure*}
\medskip
Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}. They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly~0 for very high-frequency words.
To make things clear, this value of $.33$ means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.

@@

\medskip
Results are gathered in Fig.~\ref{fig:feature-variations}.
We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
\begin{APAenumerate}
\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
\end{APAenumerate}
The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
%\rk{Cam: On vire ou bien?}

@@

>>>>

\todo{
\#19: better explain choice of single-substitution case, referring to further down in (1) substitution detection and (2) susceptibility measure
}

We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation. To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.

@@

\todo{\#17: go into more detail on how quotes are selected by Leskovec et al. (or explain there's no info if there's none)}

@@

\todo{\#17: precision/recall analysis on anti-spam and language}

@@

\todo{\#16:
(1) update the set of features,
(2) better explain feature selection: no PCA or all-feature regression (a predictive model doesn't differentiate correlated features), it's robust feature selection with domain knowledge
(3) note clustering values have changed because of weights
}

@@

\todo{\#19: show more clearly why this is more robust with single-substitution, justifying our choice of this case}

\todo{\#17: explain all models and the work involved, with graphs in annex}

@@

\medskip

\todo{\#12: fix substitution counts}

Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Fig.~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.  In practice, from the 2.43m initial occurrences spread into 45,749 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 6,172 real substitutions obeying to this model. From these substitutions we remove those featuring stop words, minor spelling changes (e.g. center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 1,051 valid substitutions.

@@

\todo{\#19: explain if we're susceptible to bias from single-substitution}

@@

%\begin{figure*}[!th]
%    \centering
%    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
%    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
%    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
%    \label{fig:feature-susceptibilities}
%\end{figure*}
\medskip

\todo{\#16:
(1) show POS don't differ,
(2) note that stopwords are excluded, which excludes POS open/closed discussion
}

\todo{\#16: explain we do this on quartiles, rework results discussion}

\todo{\#17: show other feature variations and other models in annex}

\todo{\#15: relate to missed literature}

Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}. They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly~0 for very high-frequency words.
To make things clear, this value of $.33$ means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.

@@

\medskip

\todo{\#16:
(1) add H00 lines to account for semantic similarity (nothing more fancy since it involves a threshold),
(2) rework results discussion
}

\todo{\#17: show other feature variations and other models in annex}

\todo{\#17: explain binning}

\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}

\todo{\#15: relate to missed literature}

Results are gathered in Fig.~\ref{fig:feature-variations}.
We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
\begin{APAenumerate}
\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
\end{APAenumerate}
The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
%\rk{Cam: On vire ou bien?}

@@

\todo{\#16: add the taking-context-into-account parts, potentially moving the non-relative results discussions to here:
(1) explain and add sentence-relative susceptibility by quartiles, which has a perfectly clear interpretation + results discussion,
(2) explain and add sentence-relative variations + results discussion
(3) link those results to non sentence-relative values
}

\todo{\#17: show other feature variations and other models in annex}

\todo{\#15: relate to missed literature}
